Measuring Hash Trustworthiness via Collision Utility Metrics: Logical Cryptanalysis of MD4

### 2018-02-15 dsn-2018/rebuttal-response.txt

Thank you to all of the reviewers. We have fixed the grammatical and
typographical errors in the paper and will work on better organizing the
paper to highlight our reviewers' comments.

This paper has two major parts: 1) suggesting metrics for measuring the
trustworthiness of a hash function and 2) developing new logical
cryptanalysis (LC) techniques for analyzing the structure of hash functions.
We selected MD4 not to show that it was untrustworthy, but because it was
widely agreed to be so. In the interests of reproducibility, we apply our
metrics to a known-untrustworthy function, which has multiple pre-existing
attacks against it. Further, our new LC techniques are more broadly applicable
to the entire class of Merkel-Damgård; while we have preliminary results on
MD5 and SHA-1, these are immature and not in the scope of this paper. However,
while the utility metrics apply to SHA-3 (and broadly to most standardized
hash functions), the new LC techniques will not immediately transfer. We will
update the submission to make these distinctions more clear.

To Reviewer #285E, we argue that these are five baseline metrics and not
necessarily meant to be comprehensive for all use cases. Traditionally,
the discovery of collisions in a hash function has resulted in its assumed
untrustworthiness for all use cases of that function. Metrics 1 and 2 are
independent of the target system and try to categorize how much additional
information is gained from finding a single collision (e.g., if a collision
class has many neighbors, this implies that learning about that one class
gives you multiple other candidate classes, and thus that the system is highly
structured and unsuitable for use as a cryptographically secure hash function).
Metric 3 is suitable for systems where the attacker has limited control and
most closely needs a second-preimage attack (e.g., attacking the PKI via
colliding into an existing root certificate). Metric 4 is suitable for systems
where the attacker has full control over the contents of the block, but limited
control over where that block occurs in context to other blocks. Metric 5
is most suitable to systems like source code repositories, JSON API endpoints,
or shell scripts, where only a limited character set is normal. For example,
to use a collision in SHA-1 to attack Git, metrics 4 and 5 are important, and
3 is only important if the target code needs to make sense; metrics 1 and 2
are not necessary, but useful if the proposed collision fails to achieve the
desired effect.

We are leaving aside two main features: weightings of these metrics and any
applicability of these metrics to preimage attacks, as they tend to be rare
and thus hard to study from an empirical perspective. However, we leave aside
weightings and an overarching measure of "how far is a collision from a second
preimage" for the reason that it isn't strictly necessary for every system,
and no one weighting can completely encompass all use cases for a hash
function. Thus, part of our desire to publish is to begin a dialogue with the
broader community regarding what other metrics are important, and what their
effect on the general trustworthiness of a hash is.

To Reviewer #285D, we argue that, while DSN doesn't explicitly call for
cryptographic research, our work falls into the area of
"models...for...assessing dependable systems". Because cryptographic hash
functions are well specificied and determanistic, we can model the impact of a
collision attack on a system in a SAT framework. In addition to the five
general metrics discussed above, explicitly evaluating the trustworthiness
for a specific use case is a novel advancement and well within the scope of
DSN. We present the new LC techniques, in part, to better demonstrate how we
arrived at a ~10-fold increase in collision classes and also as a basis for
a larger empirical study of the effectiveness of these metrics. Further,
part of our desire to publish in DSN is to begin a dialogue with the
broader community around dependable systems regarding what other metrics
are important to their specific use cases.

Thank you for your time.

### 2018-02-15 dsn-2018/rebuttal-response.txt

We are leaving aside two main features: weightings of these metrics and any
applicability of these metrics to preimage attacks, as they tend to be rare
and thus hard to study from an empirical perspective. However, we leave aside
weightings and an overarching measure of "how far is a collision from a second
preimage" for the reason that it isn't strictly necessary for every system,
and no one weighting can completely encompass all use cases for a hash
function. Thus, part of our desire to publish is to begin a dialogue with the
broader community regarding what other metrics are important, and what their
effect on the general trustworthiness of a hash is.

To Reviewer #285D, we argue that, while DSN doesn't explicitly call for
cryptographic research, our work falls into the area of
"models...for...assessing dependable systems". Because cryptographic hash
functions are well specified and deterministic, we can model the impact of a
collision attack on a system in a SAT framework. In addition to the five
general metrics discussed above, explicitly evaluating the trustworthiness
for a specific use case is a novel advancement and well within the scope of
DSN. We present the new LC techniques, in part, to better demonstrate how we
arrived at a ~10-fold increase in collision classes and also as a basis for
a larger empirical study of the effectiveness of these metrics. Further,
part of our desire to publish in DSN is to begin a dialogue with the
broader community around dependable systems regarding what other metrics
are important to their specific use cases.

### 2018-02-15 dsn-2018/rebuttal-response.txt

Thank you to all of the reviewers. We have fixed the grammatical and
typographical errors in the paper and will work on better organizing the
paper to highlight our reviewers' comments.

This paper has two major parts:
    1) suggesting metrics for measuring the trustworthiness of a hash function
    2) developing new logical cryptanalysis techniques (LCT) for analyzing
        the structure of hash functions.
We recognize that MD4 is untrustworthy and use it for the interests of
reproducibility. We sought to test our new metrics against a
known-untrustworthy hash which has multiple pre-existing attacks against it.
Further, our new LC techniques are more broadly applicable to the entire class
of Merkel-Damgård; while we have preliminary results on MD5 and SHA-1, these
results are immature and not in the scope of this paper. However, while the
utility metrics apply to SHA-3 (and broadly to most standardized hash
functions), the new LC techniques will not immediately transfer. We will
update the submission to make these distinctions more clear.

To Reviewer #285E, we argue that these are five baseline metrics and not
necessarily meant to be comprehensive for all use cases. Traditionally,
the discovery of collisions in a hash function has resulted in its assumed
untrustworthiness for all use cases of that function.

    - Metrics 1 and 2 are independent of the target system and try to
        quantify the information is gained from finding a single collision.
        (E.g., if a collision has a large number of neighbors, it implies
        structure in the hash function, making it untrustworthy)
    - Metric 3 is suitable for systems where the attacker has limited control
        and needs a second-preimage attack. (E.g., attacking the PKI via
        colliding into an existing root certificate).
    - Metric 4 is suitable for systems where the attacker has full control
        over the contents of the block, but limited control over where that
        block occurs in context to other blocks.
    - Metric 5 is most suitable to systems like source code repositories or
        shell scripts, where only a limited character set is normal.
For example, to use a collision in SHA-1 to attack Git, metrics 4 and 5 are
important, and 3 is only important if the target code needs to make sense;
metrics 1 and 2 are not necessary, but useful if the given collision fails.

### 2018-02-15 dsn-2018/rebuttal-response.txt

Thank you to all of the reviewers. We have fixed the grammatical and
typographical errors in the paper and will work on better organizing
the paper to highlight our reviewers' comments.

This paper has two major parts:
    1) suggesting metrics for measuring the trustworthiness of a hash function
    2) developing new logical cryptanalysis techniques (LCT) for analyzing
        the structure of hash functions.
We recognize that MD4 is untrustworthy and use it in the interest of
reproducibility. We sought to test our new metrics against a
known-untrustworthy hash which has multiple pre-existing attacks against it.
Furthermore, our new LCTs are applicable to the entire class of Merkel-Damgård,
but while we have preliminary results on MD5 and SHA-1, they are immature and
not in the scope of this paper. The utility metrics apply to SHA-3 but the new
LCTs will not immediately transfer. We will update the submission to make
these distinctions more clear.

To Reviewer #285E, we argue that these are five baseline metrics and not
necessarily meant to be comprehensive for all use cases. Traditionally,
the discovery of collisions in a hash function has resulted in its assumed
untrustworthiness for all use cases of that function; our proposed metrics
seek to address this shortcoming in the literature.
    - Metrics 1 and 2 are independent of the target system and try to
        quantify the information is gained from finding a single collision.
        (E.g., if a collision has a large number of neighbors, it implies
        structure in the hash function, making it untrustworthy)
    - Metric 3 is suitable for systems where the attacker has limited control
        and needs a second-preimage attack. (E.g., attacking the PKI via
        colliding into an existing root certificate).
    - Metric 4 is suitable for systems where the attacker has full control
        over the contents of the block, but limited control over where that
        block occurs in context to other blocks.
    - Metric 5 is most suitable to systems like source code repositories or
        shell scripts, where only a limited character set is normal.
For example, to use a collision in SHA-1 to attack Git, metrics 4 and 5 are
important, and 3 is relevant when the target code needs to make sense; 1 and
2 are not necessary, but useful if the given collision fails. Thus, the only
aspect we leave aside is how to weight these individual metrics to measure
the distance from a collision attack to a second preimage attack, but we
feel that this weighting is specific to individual use cases.

To Reviewer #285D, we argue that our work falls into DSN’s call for papers
which describe "models...for...assessing dependable systems". Because
cryptographic hash functions are well specified and deterministic, we can model
the impact of a collision attack on a system in a SAT framework. In addition to
the five general metrics discussed above, explicitly evaluating the
trustworthiness for a specific use case is a novel advancement and well within
the scope of DSN. We present the new LCTs as basis for an empirical study of
the effectiveness of these metrics and techniques. Further, part of our desire
to publish in DSN is to begin a dialogue with the broader community around
dependable systems regarding what other metrics are important to their specific
use cases.

### 2018-02-16 dsn-2018/rebuttal-response.txt

Thank you to all of the reviewers. Grammatical/typographical errors have been fixed. We will improve the organization as suggested.

We make two main contributions:

1) metrics for measuring the trustworthiness of a hash.
2) new logical cryptanalysis techniques (LCT) for analyzing the structure of hash functions.

We recognize MD4 as untrustworthy, but we use it for reproducibility. We tested our novel metrics against a known-untrustworthy hash with multiple pre-existing attacks to measure impact and improvement of our methods. Our new LCTs are applicable to the entire class of Merkel-Damgard hash functions. We have preliminary results on MD5 and SHA-1, full treatment exceeds the scope and page limit of this paper. Utility metrics apply to SHA-3, but LCTs do not trivially transfer. We will update the submission to make these distinctions clear.

#285E, we present five baseline metrics, but do not say they are comprehensive. Traditionally, the discovery of collisions in a hash function has resulted in its assumed untrustworthiness for all use cases of that function; our proposed metrics seek to address this shortcoming.

- Metrics 1 and 2 are independent of the target system and try to quantify the information gained from finding a single collision. The neighborhood density implies structure in the hash function, providing a formal measure of untrustworthiness.

- Metric 3 is suitable for systems where the attacker has limited control and needs a second-preimage attack. (E.g., attacking the PKI via colliding into a certificate).

- Metric 4 is suitable for systems where the attacker has full control over the contents of the block, but limited control over where that block occurs in context to other blocks.

- Metric 5 is most suitable to systems like source code repositories or shell scripts, where only a limited character set is normal.

For example, to use a collision in SHA-1 to attack Git, metrics 4 and 5 are important, and 3 is relevant when the target code needs to make sense; 1 and 2 are not necessary, but useful if the given collision fails. Thus, the only aspect we leave aside is how to weight these individual metrics to measure the distance from a collision attack to a second preimage attack, but we feel that this weighting is specific to individual use cases.

#285D, we argue that our work falls into DSN’s call for papers which describe "models...for...assessing dependable systems". Cryptographic hash functions are well specified and deterministic and thus we can model the impact of a collision attack on a system in a SAT framework. In addition to the five general metrics discussed above, explicitly evaluating the untrustworthiness for a specific use case is a novel advancement and well within the scope of DSN. We present the new LCTs as basis for an empirical study of the effectiveness of these metrics and techniques.

We thank all reviewers for their comments, and have incorporated them into our submission.

### 2018-03-05 dsn-2018/bare_conf.tex
DSN 2018 Submission - Rejected (#5)
Submission rejected; draft + updates.

* Initial draft of dsn-2018 paper (#3)

Authored by E. Rozier (@erozier2), A. Scheel (@cipherboy) with feedback
from K. Mayo (@katmayo7), N. De Graaf (@NathanDeGraaf),
N. Mulbrook and others.

* Updated for typos

Signed-off-by: Alexander Scheel <alexander.m.scheel@gmail.com>

* Store versioned paper

Signed-off-by: Alexander Scheel <alexander.m.scheel@gmail.com>

* Initial rebuttal

Signed-off-by: Alexander Scheel <alexander.m.scheel@gmail.com>

* Spelling

Signed-off-by: Alexander Scheel <alexander.m.scheel@gmail.com>

%\author{\IEEEauthorblockN{Alexander M. Scheel, Eric W. D. Rozier}
%\IEEEauthorblockA{Department of Computer Science\\Iowa State University of Science and Technology\\
%Ames, Iowa, USA 50011\\
%Email: \texttt{alexander.m.scheel@gmail.com, erozier@iastate.edu}}}

@@

The remainder of this paper is organized as follows. In Section
\ref{Sec:Related} we discuss previous results and show how our research
expands upon prior work. In Section \ref{Sec:Terminology}, we discuss the
terminology and notations we use throughout the remainder of the paper.
In Section \ref{Sec:Intuition}, we give the intuition behind measuring
the utility of a collision and motivate why it is useful. In Section
\ref{Sec:Logical}, we present new techniques in logical cryptanalysis to
analyze collisions in hash functions. In Section \ref{Sec:Empirical}, we
discuss the utility of existing collisions under the metrics described in
\ref{Sec:Intuition}. In Section \ref{Sec:Collision}, we discuss new
differential paths and the effectiveness of various techniques presented
in Section \ref{Sec:Logical}. Lastly, in Section \ref{Sec:Impacts}, we
discuss the broader impacts and conclusions of this research, and
conclude with Section \ref{Sec:Conclusion} discussing future work and
extensions in Section \ref{Sec:Future}.

@@

This paper draws heavily upon the public cryptanalysis of MD4, the wide
availability of high quality SAT solvers, and previous work in logical
cryptanalysis. In particular, we look at ways that SAT solving can aid
cryptanalysis directly, where as in the literature cryptanalysis is
primarily used to benchmark SAT
solvers.  A critical assumption of our use of SAT solvers as a means
to discover structures in hash functions which lead to untrustworthy
behavior is the insight that while SAT is, in general, a hard problem
for many specific and practical instances current SAT-solving
technology is quite efficient \cite{vardi2010p}. We
then apply them to aid the understanding of attacks against real world
systems, and propose general metrics that target different types of systems.
Towards this, we look at prior work in two main areas: differential and logical
cryptanalysis.

@@

In order to define our define our metrics using logical cryptanalysis
we first define a set of formal terminology and notation whn
discussing our work, and the notation of collisions within a hash.
We make the general assumption that ordered tuples and strings are
indexable via square brackets. We utilize subscripts when referring to
elements of tuples with named members, as opposed to just indexable
members.  When referring to the MD4 hash function, we use MD4 as
specified in RFC 1320 \cite{rfc1320}.

@@

We introduce a distance function, $\delta$, between collision classes by the
number of differences in intermediate rounds deltas. That is, given two
collision classes, $C_1, C_2 \in \mathcal{C}$:
\begin{align}
    \delta(C_1, C_2) = \big| \{ i : C_1[i] \neq C_2[i] \} \big|
\end{align}
We extend this distance function to include an \textit{absolute} distance,
measuring the number of intermediate rounds with non-zero differences.
We notate this as $\delta(C_1)$.

@@

Furthermore this condition extends to $\delta = d$ for some small
value of $d$ as determined by computational resources such that $d$
becomes a trustworthiness metric under constraints of current
3-CNF-SAT solving capabilities within reasonable time bounds.

@@

Evaluating flexibility becomes important as it serves as a
contrapositive to trustworthiness.  A highly flexible attack implies a
larger set of families of collision classes, and decreasing the trust
that should be placed in a hash function. To evaluate how much flexibility exists in a collision,
we propose the following metrics for evaluating the utility of a collision
class:
\begin{enumerate}
    \item The number of unique differentials a collision class has.
    \item The number of unit distance neighbors a collision class has.
    \item The maximum count of zeros in a the binary representation of a
        colliding block (and likewise with ones).
    \item Whether there exists a block which collides under multiple initial
        values.
    \item Whether or not zero, one, or both of the blocks in a collision may
        be of ASCII values under any input block difference.
\end{enumerate}
Note that the first three are quantitative measures providing some measure of
flexibility of a collision class, whereas the latter two are merely looking
for a single witness having such a property. Depending on the scenario,
specific properties of a collision may be of more interest than others.

@@

We justify our the distance heuristic, $\delta$, defined in Section
\ref{terminology:distance} in the existing literature on MD4.  Using
H. Dobbertin's \cite{Dobbertin1998} and P. Kasselman's \cite{KasselmanMD4}
collision classes, we find they have distance $\delta(C_{Dobbertin}, C_{Kasselman}) = 1$
under our metric. Since others have shown that MD4 is untrustworthy,
we hypothesized (correctly) that we could
limit our search space to small distances, exploiting the structure of MD4. This
reduction in search space decreases the time to solve a single 3-CNF-SAT
instance at the expense of restricting ourselves to highly connected, highly
similar differential paths which none-the-less yield rich and
interesting sets of collisions. We further improved our performability
by exploiting the parallalizability of this problem
noting that this problem is embarrassingly parallel: allowing us to
check each
candidate collision class independently on separate machines.

@@

The highly connected, highly similar differential paths are highlighted in
Figure \ref{Fig:Graph-Clique}: when the neighbors of H. Dobbertin's and
P. Kasselman's collision classes from Figure \ref{Fig:Graph-Intro} are viewed
independently, the large cliques become apparent. Depending on the search
strategies utilized, cliques may be expanded before new neighbors, or a
significant portion of a vertices neighbors might belong to existing cliques,
resulting in slow progress towards any external goal.

@@

In this section we present our empirical results, evaluating the
utility of existing published collisions by exploring their structure
with 3-CNF-SAT, and building representative neighborhoods using our
distance metrics.  We present these results in the form of an
exploration of unique differentials, exploration of unit distance
neighborhoods, maximum count of zeros in the binary representation of
a colliding block, the existance of blocks which collide under
multiple initial values, and ASCII block collisions as discussed
previously in Section \ref{Sec:Intuition}.

@@

We present results in Table \ref{table:zeroes} which represent the maximum count of zeroes in a block
for existing collisions. Note that X. Wang's attack is the leader, with 509
maximum zero bits in the collision pair. Further, while Y. Sasaki improved
the bound in other areas, this was not one of them. Of note too,
H. Dobbertin's and P. Kasselman's collisions are essentially identical in
this metric.

\begin{table}
    \caption{Number of Zeroes in a Colliding Block}
    \label{table:zeroes}
    \begin{tabular}{c c c c}
        \textbf{Attack} & \textbf{Count} & \textbf{Attack} & \textbf{Count} \\
        X. Wang's & 509 & Y. Sasaki's & 494 \\
        H. Dobbertin's & 504 & P. Kasselman's & 504 \\
        M. Schl{\"a}ffer's & 506 & & \\
    \end{tabular}
\end{table}

@@

Through the exploration of the neighborhood technique, we have found over
35,000 new differential paths which originated from the five original published
collisions in MD4. This shows that the neighborhood definition is sufficient
to produce a large quantity of new collision classes. Further, we give
evidence for propagating neighborhoods across reduced round versions of MD4.

@@

We noted in table \ref{table:neighborhood} that the unit distance neighborhood of
the existing attacks were non-empty. By repeatedly expanding neighborhoods of
known collision classes, we can construct paths in the collision space. We
made the following observations:

@@

We demonstrate the utility of five metrics of the flexibility present
in a collision on the basis of 1) the number of unique differentials
within a collision class; 2) the size of the unit distance
neighborhood for a collision class; 3) the maximum count of zeros in a
binary representation of a colliding block; 4) the existance of blocks
which collide under multiple initial values; and 5) the presence of
ASCII valued inplut blocks within a collision.  These metrics proved
capable when evaluating the flexibility of collisions in the published
literature, and showed that flexibility is a multi-dimensional
quantity that must be compared over multiple distance metrics, as not all
collisions dominate universally.

@@

In order to help the community build upon or work, and for the
purposes of reproducability and ascertainment of correctness the
research, data, and all associated code are provided on an open access
and open source basis under the GNU General Public License Version 3
\footnote{\url{https://www.gnu.org/licenses/gpl-3.0.en.html}}.
The code itself is hosted on GitHub at \url{https://github.com/XXXXXXXXXXXXXXXXXX}.
For additional details or to request data associated with the project, please
contact the authors.

>>>>

\author{\IEEEauthorblockN{Alexander M. Scheel, Eric W. D. Rozier}
\IEEEauthorblockA{Department of Computer Science\\Iowa State University of Science and Technology\\
Ames, Iowa, USA 50011\\
Email: \texttt{alexander.m.scheel@gmail.com, erozier@iastate.edu}}}

@@

The remainder of this paper is organized as follows. In Section
\ref{Sec:Related} we discuss previous results and show how our research
expands upon prior work. In Section \ref{Sec:Terminology}, we discuss the
terminology and notations we use throughout the remainder of the paper.
In Section \ref{Sec:Intuition}, we give the intuition behind measuring
the utility of a collision and motivate why it is useful. In Section
\ref{Sec:Logical}, we present new techniques in logical cryptanalysis to
analyze collisions in hash functions. In Section \ref{Sec:Empirical}, we
discuss the utility of existing collisions under the metrics described in
\ref{Sec:Intuition}. In Section \ref{Sec:Collision}, we discuss new
differential paths and the effectiveness of various techniques presented
in Section \ref{Sec:Logical}. Lastly, in Section \ref{Sec:Conclusion}, we
discuss the broader impacts and conclusions of this research, and
conclude by discussing future work and extensions in Section \ref{Sec:Future}.

@@

This paper draws heavily upon the public cryptanalysis of MD4, the wide
availability of high quality SAT solvers, and previous work in logical
cryptanalysis. In particular, we look at ways that SAT solving can aid
cryptanalysis directly, where as in the literature cryptanalysis is
primarily used to benchmark SAT
solvers.  A critical assumption of our use of SAT solvers as a means
to discover structures in hash functions which lead to untrustworthy
behavior is the insight that while SAT is, in general, a hard problem
for many specific and practical instances, current SAT-solving
technology is quite efficient \cite{vardi2010p}. We
then apply them to aid the understanding of attacks against real world
systems, and propose general metrics that target different types of systems.
Towards this, we look at prior work in two main areas: differential and logical
cryptanalysis.

@@

In order to define our metrics using logical cryptanalysis
we first define a set of formal terminology and notation whn
discussing our work, and the notation of collisions within a hash.
We make the general assumption that ordered tuples and strings are
indexable via square brackets. We utilize subscripts when referring to
elements of tuples with named members, as opposed to just indexable
members.  When referring to the MD4 hash function, we use MD4 as
specified in RFC 1320 \cite{rfc1320}.

@@

We introduce a distance function, $\delta$, between collision classes by the
number of differences in intermediate rounds deltas. That is, given two
collision classes, $C_1, C_2 \in \mathcal{C}$:
\begin{align}
    \delta(C_1, C_2) = \big| \{ i : C_1[i] \neq C_2[i] \} \big|
\end{align}
We extend this distance function to include an \textit{absolute} distance,
measuring the number of intermediate rounds with non-zero differences.
We denote this as $\delta(C_1)$.

@@

Furthermore, this condition extends to $\delta = d$ for some small
value of $d$ as determined by computational resources such that $d$
becomes a trustworthiness metric under constraints of current
3-CNF-SAT solving capabilities within reasonable time bounds.

@@

Evaluating flexibility becomes important as it serves as a
contrapositive to trustworthiness.  A highly flexible attack implies a
larger set of families of collision classes, and decreasing the trust
that should be placed in a hash function. To evaluate how much flexibility exists in a collision,
we propose the following metrics for evaluating the utility of a collision
class:
\begin{enumerate}
    \item The number of unique differentials a collision class has.
    \item The number of unit distance neighbors a collision class has.
    \item The maximum count of zeros in the binary representation of a
        colliding block (and likewise with ones).
    \item Whether there exists a block which collides under multiple initial
        values.
    \item Whether or not zero, one, or both of the blocks in a collision may
        be of ASCII values under any input block difference.
\end{enumerate}
Note that the first three are quantitative measures providing some measure of
flexibility of a collision class, whereas the latter two are merely looking
for a single witness having such a property. Depending on the scenario,
specific properties of a collision may be of more interest than others.

@@

We justify our distance heuristic, $\delta$, defined in Section
\ref{terminology:distance} in the existing literature on MD4.  Using
H. Dobbertin's \cite{Dobbertin1998} and P. Kasselman's \cite{KasselmanMD4}
collision classes, we find they have distance $\delta(C_{Dobbertin}, C_{Kasselman}) = 1$
under our metric. Since others have shown that MD4 is untrustworthy,
we hypothesized (correctly) that we could
limit our search space to small distances, exploiting the structure of MD4. This
reduction in search space decreases the time to solve a single 3-CNF-SAT
instance at the expense of restricting ourselves to highly connected, highly
similar differential paths which none-the-less yield rich and
interesting sets of collisions. We further improved our performability
by exploiting the parallalizability of this problem
noting that this problem is embarrassingly parallel: allowing us to
check each
candidate collision class independently on separate machines.

@@

The highly connected, highly similar differential paths are highlighted in
Figure \ref{Fig:Graph-Clique}: when the neighbors of H. Dobbertin's and
P. Kasselman's collision classes from Figure \ref{Fig:Graph-Intro} are viewed
independently, the large cliques become apparent. Depending on the search
strategies utilized, cliques may be expanded before new neighbors, or a
significant portion of a vertex's neighbors might belong to existing cliques,
resulting in slow progress towards any external goal.

@@

In this section we present our empirical results, evaluating the
utility of existing published collisions by exploring their structure
with 3-CNF-SAT, and building representative neighborhoods using our
distance metrics.  We present these results in the form of an
exploration of unique differentials, exploration of unit distance
neighborhoods, maximum count of zeros in the binary representation of
a colliding block, the existence of blocks which collide under
multiple initial values, and ASCII block collisions as discussed
previously in Section \ref{Sec:Intuition}.

@@

We present results in Table \ref{table:zeros} which represent the maximum count of zeros in a block
for existing collisions. Note that X. Wang's attack is the leader, with 509
maximum zero bits in the collision pair. Further, while Y. Sasaki improved
the bound in other areas, this was not one of them. Of note too,
H. Dobbertin's and P. Kasselman's collisions are essentially identical in
this metric.

\begin{table}
    \caption{Number of Zeroes in a Colliding Block}
    \label{table:zeros}
    \begin{tabular}{c c c c}
        \textbf{Attack} & \textbf{Count} & \textbf{Attack} & \textbf{Count} \\
        X. Wang's & 509 & Y. Sasaki's & 494 \\
        H. Dobbertin's & 504 & P. Kasselman's & 504 \\
        M. Schl{\"a}ffer's & 506 & & \\
    \end{tabular}
\end{table}

@@

Through the exploration of the neighborhood technique, we have found over
35,000 new differential paths which originated from the five originally published
collisions in MD4. This shows that the neighborhood definition is sufficient
to produce a large quantity of new collision classes. Further, we give
evidence for propagating neighborhoods across reduced round versions of MD4.

@@

We noted in Table \ref{table:neighborhood} that the unit distance neighborhood of
the existing attacks were non-empty. By repeatedly expanding neighborhoods of
known collision classes, we can construct paths in the collision space. We
made the following observations:

@@

We demonstrate the utility of five metrics of the flexibility present
in a collision on the basis of 1) the number of unique differentials
within a collision class; 2) the size of the unit distance
neighborhood for a collision class; 3) the maximum count of zeros in a
binary representation of a colliding block; 4) the existence of blocks
which collide under multiple initial values; and 5) the presence of
ASCII valued input blocks within a collision.  These metrics proved
capable when evaluating the flexibility of collisions in the published
literature, and showed that flexibility is a multi-dimensional
quantity that must be compared over multiple distance metrics, as not all
collisions dominate universally.

@@

In order to help the community build upon our work, and for the
purposes of reproducability and ascertainment of correctness, the
research, data, and all associated code are provided on an open access
and open source basis under the GNU General Public License Version 3
\footnote{\url{https://www.gnu.org/licenses/gpl-3.0.en.html}}.
The code itself is hosted on GitHub at \url{https://github.com/XXXXXXXXXXXXXXXXXX}.
For additional details or to request data associated with the project, please
contact the authors.
